---
title: "Data Wrangling"
output: pdf_document
date: "2024-09-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r}
contestant_data <- read_csv("/home/rstudio/work/data/survivor/contestant_table.csv")
season_data <- read_csv("/home/rstudio/work/data/survivor/season_table.csv")
tribe_data <- read_csv("/home/rstudio/work/data/survivor/tribe_table.csv")
```

I think I'll def use contestant_data and maybe season_data too but not sure about tribe_data

```{r}
## probably should look at proportions for most of these 
ggplot(contestant_data, aes(x=age)) +
  geom_histogram()

ggplot(contestant_data, aes(x=gender)) +
  geom_bar()

ggplot(contestant_data, aes(x=poc)) +
  geom_bar()


contestant_data %>%
  group_by(profession) %>%
  tally()%>% arrange(desc(n))

contestant_data %>%
  group_by(state) %>%
  tally()%>% arrange(desc(n))
```
I think the profession needs to be tidied:
Okay what can i combine?
- Teacher: 4th Grade Teacher, Art Teacher, Elementary Teacher, English Teacher, Gym Teacher, high school teacher
- Professor: English Professor, English professor
- Actor: Actor, Actress, Aspiring Actress, former tv teen star
- Veteran: Air Force Veteran, Army Veteran, Iraq War Veteran, marine
- Student: Art Student, Business Student, College Student, Cosmetology Student, Crossfit Coach/Physical Therapy Student. Dental Student, Economics Student, full-time student/sales, grad student, graduate student, harvard law student, high school student, Ivy League Student, law school student, law student, law studnet/former military, mba student, medical student
- Career Consultant + Career Counselor
- Coach: Baseball/Dating Coach, Basketball Coach, Cheerleading Coach, Crossfit Trainer, Crossfit Coach/Project Coordinator, fitness instructor, former nfl coach, gymnastics coach, hiking guide, life coach/speaker/yoga teacher, navy swim instructro
- Gym Owner: Competitive Cheerleading Gym Owner, gym owner
- Attorney: Attorney + Civil Rights Attorney + Defense Attorney, Deputy District Attorney, Divorce Attorney, Attorney/Former NFL Player + Attorney/Retired NFL Player
- Barista: Barista + Coffee Barista
- Computer Programmer: Computer Programmer + Computer Engineer
- Construction: Construction Company Vice President, Construction Supervisor, Construction Worker, highway construction
- Designer + Aspiring Designer
- Doctor: Doctor, ER Doctor
- Entertainment Host + Writer: Entertainment Host and Writer, Entertainment Host/Writer
- Firefighter: Firefighter, Fireman/EMT
- Fisher: fisherwoman, fishing boat captain, fishing guide, fishmonnger
- Pro Athlete: former nba all-star, former nba player, former mlb player, former mfl quarterback, former nhl player,former professional cyclist, former pro cyclist/shop manager, mixed martial arts fighter, nfl player
- writer: aspiring writer, freelance writer
- farmer: goat farmer, farmer, goat racher
- sales: elevator salesman, graphic saleswoman, insurance saleswoman
- Hairdresser: hairdresser, hairsytlist
- Singer: country music signer, former pop star, heavy metal musican
- Animal Trainer: horse trainer, dolphin trainer, dog trainer
- Internet Project Manager: Internet Projects Manager, Internet Project Manager
- Jewelry: Jeweler, Jewelry Designer, Jewelry Salesman/Photographer
- lifeguard: lifeguard + former lifeguard


** i stiopped on page 32 -- idk if i want to keep doing this
** note:  Cheif Lifegaurd, industrial engineer

```{r}

## winners
contestant_data %>%
  filter(finish==1) %>%
  ggplot(aes(x=age))+
  geom_histogram()

contestant_data %>%
  filter(finish==1) %>%
  ggplot(aes(gender)) +
 geom_bar(aes(fill = as.factor(poc)), position = "dodge")

contestant_data %>%
  filter(finish==1) %>%
  ggplot(aes(poc)) +
  geom_bar()

contestant_data %>%
  filter(finish==1) %>% 
  group_by(profession) %>% tally() %>% arrange(desc(n))

contestant_data %>%
  filter(ftc==1)

```

```{r}
## first voted out
contestant_data %>%
  filter(
    finish==16 & num_season %in% c(1,2,3,4,5,6,7,12,15,18) |
    finish==18 & num_season %in% c(8,9,11,17,22,23,24,25,28,29,30,32,35,38,41,42,43) |
    finish==19 & num_season ==14 |
    finish==20
  )  %>%
  ggplot(aes(x=age)) +
  geom_histogram(aes(fill = as.factor(gender)))

contestant_data %>%
  filter(
    finish==16 & num_season %in% c(1,2,3,4,5,6,7,12,15,18) |
    finish==18 & num_season %in% c(8,9,11,17,22,23,24,25,28,29,30,32,35,38,41,42,43) |
    finish==19 & num_season ==14 |
    finish==20
  )  %>%
  ggplot(aes(gender)) +
 geom_bar(aes(fill = as.factor(poc)), position = "dodge")

contestant_data %>%
  filter(
    finish==16 & num_season %in% c(1,2,3,4,5,6,7,12,15,18) |
    finish==18 & num_season %in% c(8,9,11,17,22,23,24,25,28,29,30,32,35,38,41,42,43) |
    finish==19 & num_season ==14 |
    finish==20
  )  %>%
  ggplot(aes(poc)) +
  geom_bar()
```

```{r}
# made the merge
contestant_data %>%
  filter(merge==1) %>%
  ggplot(aes(age)) +
  geom_histogram()

contestant_data %>%
  filter(merge==1) %>%
  ggplot(aes(gender)) +
  geom_bar()

contestant_data %>%
  filter(merge==1)%>%
  group_by(profession) %>% tally() %>% arrange(desc(n))
```

What are some things I could look at?

What types of people win? What types of people make final tribal (+ dont win?) What types of people get voted out first? What types of people don't make the merge v do make the merge? Is there a "lucky" buff color -- i.e. the first color buff that a contest is assigned to - compare this with winners 1st buff colors.


** okay wait... apparently there is an R package called `survivoR` that has data on different things of each season!!!! Im gunna go through them now:

```{r}
advantage_details
advantage_data <- advantage_movement
castaways
castaway_details
challenge_description
challenge_results
challenge_summary
confessionals %>% arrange(desc(confessional_count))
episodes %>% arrange(desc(viewers))
screen_time
season_summary
vote_history

confessional_data <- confessionals
```

```{r}
challenge_data <- challenge_summary %>%
 filter(str_starts(version_season, "US"))

## I could calculate win percentage by adding up the number of wins and dividing by the number of challenges they competed in

```


Okay... plan of action -- this is too much data and its overwhelming so I'm gunna stick with what I have except I want to add win percentage. So i can calculate the win percentages and then filter that data so that it matches with the data I have (bc i only have until season 43 i think -- or I can add the contestants that I'm missing bc I do have all their info). Then i can merge it into my contestant dataset so I can see how win percentage plays a part in your overall play. If im feeling up to it i could also go through the advantage data and make a boolean for if a contestant found an advantage (and if they used it -- used it correctly -- and number of advantages found)
